---
title: "Time Series Lab 2"
author: "Randall Plyler"
date: "11/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("C:/Users/randa/Dropbox/Masters/Autumn/TBANLT 540 Applied Regression Analysis/Time Series Lab 2"))
```


Library Packages

```{r library}

library(tidyverse)
library(readxl)
library(corrgram)
library(tseries)
library(forecast)
```


```{r }
RainSanDiegoAll <- read_csv("2794315.csv")

```

```{r data structure}
dim(RainSanDiegoAll)    #returns the dimensions of an object
str(RainSanDiegoAll)
sum(is.na(RainSanDiegoAll)) #returns how many observations have "na"
mae<-function(error) {mean(abs(error))}

```


```{r}

view(RainSanDiegoAll)
#RainSanDiegoAll$PRCP <- as.character(as.logical(RainSanDiegoAll$PRCP))
#RainSanDiegoAll$DATE <- as.character(as.logical(RainSanDiegoAll$DATE))
#RainSanDiegoAll$ACMH <- as.character(as.logical(RainSanDiegoAll$ACMH))
#RainSanDiegoAll$ACSH <- as.character(as.logical(RainSanDiegoAll$ACSH))
#RainSanDiegoAll$WDFG <- as.character(as.logical(RainSanDiegoAll$WDFG))
#RainSanDiegoAll$WDFM <- as.character(as.logical(RainSanDiegoAll$WDFM))
#RainSanDiegoAll$WSF1 <- as.character(as.logical(RainSanDiegoAll$WSF1))
#RainSanDiegoAll$WSFG <- as.character(as.logical(RainSanDiegoAll$WSFG))
#RainSanDiegoAll$WSFM <- as.character(as.logical(RainSanDiegoAll$WSFM))
#RainSanDiegoAll$WT07 <- as.character(as.logical(RainSanDiegoAll$WT07))
#RainSanDiegoAll$WT09 <- as.character(as.logical(RainSanDiegoAll$WT09))
#RainSanDiegoAll$WT11 <- as.character(as.logical(RainSanDiegoAll$WT11))
#RainSanDiegoAll$WT15 <- as.character(as.logical(RainSanDiegoAll$WT15))
#RainSanDiegoAll$WV01 <- as.character(as.logical(RainSanDiegoAll$WV01))
#RainSanDiegoAll$WV20 <- as.character(as.logical(RainSanDiegoAll$WV20))


#RainSanDiegoAll$SNOW [is.na(RainSanDiegoAll$SNOW )] <- '0' 
#RainSanDiegoAll$ACMH [is.na(RainSanDiegoAll$ACMH )] <- '0' 
#RainSanDiegoAll$ACSH [is.na(RainSanDiegoAll$ACSH )] <- '0' 
#RainSanDiegoAll$WDFG [is.na(RainSanDiegoAll$WDFG )] <- '0' 
#RainSanDiegoAll$WDFM [is.na(RainSanDiegoAll$WDFM )] <- '0' 
#RainSanDiegoAll$WSF1 [is.na(RainSanDiegoAll$WSF1 )] <- '0' 
#RainSanDiegoAll$WSFG [is.na(RainSanDiegoAll$WSFG )] <- '0' 
#RainSanDiegoAll$WSFM [is.na(RainSanDiegoAll$WSFM )] <- '0' 
#RainSanDiegoAll$WT07 [is.na(RainSanDiegoAll$WT07 )] <- '0' 
#RainSanDiegoAll$WT09 [is.na(RainSanDiegoAll$WT09 )] <- '0' 
#RainSanDiegoAll$WT11 [is.na(RainSanDiegoAll$WT11 )] <- '0' 
#RainSanDiegoAll$WT15 [is.na(RainSanDiegoAll$WT15 )] <- '0' 
#RainSanDiegoAll$WV01 [is.na(RainSanDiegoAll$WV01 )] <- '0' 
#RainSanDiegoAll$WV20 [is.na(RainSanDiegoAll$WV20 )] <- '0' 



#sum(is.na(RainSanDiegoAll))

#RainSanDiegoAll[is.na(RainSanDiegoAll)] <- 0 #replaces "na" with 0. This is a choice, statistically, but you can't run the regression without it

#sum(is.na(RainSanDiegoAll))


```

Create a time series dataset for decomposition #1 #2

```{r time_series_creation}

RainSanDiegoAll$DATE<-as.Date(RainSanDiegoAll$DATE, format="%m/%d/%Y")
RainSanDiegoAll$PRCP<-as.numeric(RainSanDiegoAll$PRCP)
view(RainSanDiegoAll)
MonthlyRain<-aggregate(list(rain = RainSeattleAll$PRCP), 
          list(month = cut(RainSanDiegoAll$DATE, "month")), 
          mean)
MonthlyRain
View(MothlyRain)
MonthlyRain2<-ts(RainSanDiegoAll$PRCP, frequency = 12, start = c(1940,1))
Rain<-ts(RainSanDiegoAll$PRCP, frequency = 365, start = c(1940,1))
adf.test(Rain)
adf.test(MonthlyRain2)

```


Create a plot of the time series


```{r plots}
plot.ts(Rain)
plot.ts(MonthlyRain2)

```


Identify the trend/season/random components #3

```{r decomposition}
RainParts<-decompose(Rain)
RainMonthParts<-decompose(MonthlyRain2)
plot(RainParts)
plot(RainMonthParts)
```

Modeling using exponential smoothing - Full data #4

```{r HoltWintersFull}
RainModel1<-HoltWinters(Rain)
RainModel1
RainModel1$SSE
plot(RainModel1, col=4, col.predicted=6)
residualsHolt1<-residuals(RainModel1)
plot(residualsHolt1)
acf(residualsHolt1)
pacf(residualsHolt1)
mae(residualsHolt1)
```

Modeling using exponential smoothing - Monthly data #5

```{r HoltWintersMonthly}
RainModel2<-HoltWinters(MonthlyRain2)
RainModel2
RainModel2$SSE
plot(RainModel2, col=4, col.predicted=6)
residualsHolt2<-residuals(RainModel2)
plot(residualsHolt2)
acf(residualsHolt2)
pacf(residualsHolt2)

```


Forecasting using exponential smoothing - Full Data

```{r HWForecastFull}
RainForecast<-forecast(Rain, h=400)
plot(RainForecast)
```

Forecasting using exponential smoothing - Monthly Data

```{r HWForecastFull}
RainForecast2<-forecast(MonthlyRain2, h=100)
plot(RainForecast2)
```

Modeling using an auto.arima model - Full Data #6

```{r AutoArimaFull}
par(mfrow=c(1,2))
acf(Rain)
pacf(Rain)


RainArima<-auto.arima(Rain, seasonal = TRUE)
RainArima
acf(ts(RainArima$residuals), main='ACF Residual - Full')
pacf(ts(RainArima$residuals), main='PACF Residual - Full')

```


Modeling using an auto.arima model - Monthly Data #7

```{r AutoArimaMonthly}
acf(MonthlyRain2)
pacf(MonthlyRain2)

RainArima2<-auto.arima(MonthlyRain2, seasonal = TRUE)
RainArima2

acf(ts(RainArima2$residuals), main='ACF Residual - Monthly')
pacf(ts(RainArima2$residuals), main='PACF Residual- Monthly')

```


```{r predictions}
RainArima3<-Arima(MonthlyRain2, order = c(5, 1, 0), seasonal = c(5, 0, 0),
                       include.mean = FALSE, include.drift = FALSE, include.constant =TRUE,
                       method = "ML")

RainArima3

RainArima4<-Arima(MonthlyRain2, order = c(5, 1, 0), seasonal = c(4, 0, 0),
                  include.mean = FALSE, include.drift = FALSE, include.constant =TRUE,
                  method = "ML")
RainArima4



prediction1=predict(RainArima,n.ahead=100)
prediction1
plot(forecast(RainArima,h=500))

prediction2=predict(RainArima2,n.ahead=10)
prediction2
plot(forecast(RainArima2,h=100))

prediction3=forecast(RainArima3)
prediction3
plot(forecast(RainArima3,h=100))

prediction4=forecast(RainArima4)
prediction4
plot(forecast(RainArima4,h=10))


```


```{r predictions}

mae(residualsHolt1)
mae(residualsHolt2)
mae(RainForecast$residuals)
mae(RainForecast2$residuals)
mae(RainArima$residuals)
mae(RainArima2$residuals)
mae(RainArima3$residuals)
mae(RainArima4$residuals)

```


